---
title: "A3: Incarceration"
output: html_document
---

### Introduction

The problem with the US prison system is the disparity between black and white people in the system. Specifically, black people are incarcerated at  high rates compared to other groups due to the ongoing racism. In this assignment I will explore data regarding female imprisonment and female demographics in prison (race). The variables I will be exploring are the female population, female demographics, and the distribution of female imprisonment in US states. I want to explore specifically female because as a female, I want to explore the situation of female imprisonment and to get a chance to analyze these variables. 

### Summary Information

The 5 values calculated are the which state with the highest ratio of Black to White female prison population, which state has the highest total female prison population in 2018, which state has the highest total Black female prison population in 2018, what year has the highest total female prison population, and average Black female prison population throughout the years. What I found was the state with the highest ratio of black to white female prison population was New Hampshire, the state with the highest female prison population was Arizona, and the state with the highest black female population was Texas. I also found the year with the highest female prison population was surprisingly in 1998. Lastly, I found the average Black female prison population throughout the years. The average lines up with the other calculations. The average of black prison population was quite high around year 1998 which was the time where it was the highest population of female imprisonment. 

1. Find the state with the highest ratio of Black to White female prison population: New Hampshire
2. Which state has the highest total female prison population in 2018: Arizona
3. Which state has the highest total Black female prison population in 2018: Texas
4. Which year has the highest total female prison population: 1998
5. Average Black female prison population throughout the years: 
```{r, echo = FALSE, message=FALSE, warning=FALSE, code = readLines("table.R")}
```

### The Dataset

The institute that collected the prison data alongside other datasets is called the Vera Institute of Justice. The primary source of prison data came from the National Corrections Reporting Program (NCRP), which was created by the Bureau of Justice Statistics (BJS) in 1983. The reasons why the data was collected were because Vera stated this topic is often overlooked in national conversations about mass incarceration, and not many sources provide a focus on the data and stats on incarceration. The dataset I used was “Dataset about prison population per county/state”. In this dataset there were 153811 rows and 36 columns. One ethical question I have is does this dataset fully represent the information on the incarceration of the US. What I mean by this is that there are many other groups that aren’t included in this dataset. People who are trans, intersex, and others that don’t fit into the category of female or male. What if the people who are imprisoned and don’t identify as male or female... do they get left out... excluded? This ties in the limitation of this dataset. I would argue that it doesn’t provide an essential demographic option to understand the incarceration problem in the US deeper. 

### Trends Over Time Chart
```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("chart1_example.R")
line_chart2 
```
Here is a chart displaying the distribution of the race of the female prison population in Washington State throughout the years. It is surprising that the White female population was more that the Black female population. From other calculations, other states are different. It is a good insight on the pattern of the demographic in Washington State. 
### Variable Comparison Chart
```{r, echo=FALSE, message=FALSE, warning=FALSE} 
source("chart2_example.R")
scatterplot
```
This chart is displaying the total female prison population in Washington State throughout the years. It is interesting to not that there has been a steady incline since the 1990 and around 2003 the population is slowly inclining. 
### Map
```{r, echo=FALSE, message=FALSE, warning=FALSE} 
source("map_example.R")
map
```
This map of the US states are showcasing the distribution of the female prison population in 2018. Texas, California, and Florida seems to have the most female prison population. 
